---
title: "Rmarkdown_dataTidying"
author: "Patrick"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}
library(dplyr) #Include the library for dplyr and tidyr.
library(tidyr)

```
Note: warning messages indicate that function names in this package overlap with function names from another package.  

Set warning = F and message = F to suppress messages.

To use a specific function from another package, use  `package::function(...)`

# Data cleaning

Read in data file.

```{r}

catch <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(catch)
```

Note: cmd + shift + m gives %<%, the pipe operator.
Note: for pipe, first argument comes from previous output in pipe.

* First "select" the data that is actually needed: Remove marginal sum and notes column.
* Then "gather" some columns together: move from wide to long format.  
Gather columns into a new heading "species" ... each row will have a species property and also another property, catch.  ("Spread" is the opposite.)

```{r}

#Clean up data.
# catch_cleaned <- catch_df %>% 
#   select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) or use - to remove columns.
catch_long <- catch %>% 
  select(-All, -notesRegCode) %>% 
  gather(key = "species", value = "catch", -Year, -Region)

head(catch_long)

```

Update column name for catch to reflect the units.
Also we want to update the values in the catch column so we can change the units.

```{r}
  catch_cleaned <- catch_long %>% 
    rename(catch_thousands = catch) #%>% 
    #mutate(catch = catch_thousands * 1000)
  head(catch_cleaned)
```

```{r}
#Testing chunk, to fix catch column being a string.
test_catch <- as.integer(catch_cleaned$catch_thousands)  #Convert strings to integers.
i <- which(is.na(test_catch)==1) #Find where there are no numbers.
catch_cleaned[i,]   #,] means everything in that dimension
```


# Split-Apply-Combine

# Joins