---
title: "Rmarkdown_dataTidying"
author: "Patrick"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}
library(dplyr) #Include the library for dplyr and tidyr.
library(tidyr)

```
Note: warning messages indicate that function names in this package overlap with function names from another package.  

Set warning = F and message = F to suppress messages.

To use a specific function from another package, use  `package::function(...)`

# Data cleaning

Read in data file.

```{r}

catch <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)

head(catch)
```

Note: cmd + shift + m gives %<%, the pipe operator.
Note: for pipe, first argument comes from previous output in pipe.

* First "select" the data that is actually needed: Remove marginal sum and notes column.
* Then "gather" some columns together: move from wide to long format.  
Gather columns into a new heading "species" ... each row will have a species property and also another property, catch.  ("Spread" is the opposite.)

```{r}

#Clean up data.
# catch_cleaned <- catch_df %>% 
#   select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) or use - to remove columns.
catch_long <- catch %>% 
  select(-All, -notesRegCode) %>% 
  gather(key = "species", value = "catch", -Year, -Region)

head(catch_long)

```

Update column name for catch to reflect the units.
Also we want to update the values in the catch column so we can change the units.
Due to an erroneous value due to OCR (optical character recognition) issue, the value "I" was changed to 1 for one entry.

```{r}
  catch_cleaned <- catch_long %>% 
    rename(catch_thousands = catch) %>% 
    mutate(catch_thousands = ifelse(catch_thousands=="I","1",catch_thousands)) %>% 
    mutate(catch_thousands = as.integer(catch_thousands)) %>% 
    mutate(catch = catch_thousands * 1000)
  head(catch_cleaned)
```

```{r, eval=F, echo=F }
#Testing chunk, to fix catch column being a string.
test_catch <- as.integer(catch_cleaned$catch_thousands)  #Convert strings to integers.
i <- which(is.na(test_catch)==1) #Find where there are no numbers.
catch_cleaned[i,]   #,] means everything in that dimension
```


# Split-Apply-Combine

Calculate the total catch by region.

```{r}
  catch_total <- catch_cleaned %>% 
    group_by(Region, Year) %>% 
    summarize(catch_region = sum(catch),
              n_obs=n())
  catch_total
  
```

Applying a filter to our data to calculate the statistics only for Chinook Salmon.
Also sort according to a specified value, e.g. Year

```{r}
  catch_chinook <- catch_cleaned %>% 
    filter(species == "Chinook" & Year > 1990 & Region == "SSE") %>% 
    arrange(-Year)
  head(catch_chinook)
```

Summary of useful functions for manipulating data:

*select
*filter
*mutate
*rename
*gather
*spread
*arrange
*separate
*unite

# Joins

Here we will use a "left join" to join the region definitions dataframe to the catch dataframe.
The key that joins these tables together is the region.

First we'll load the region dataframe.

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)
head(region_defs)

```

Then we'll clean up the data frame using the "select" function.

```{r}
  region_clean <- region_defs %>% 
    select(code, mgmtArea) %>% 
    rename(Region = code)
```

Then we'll join the catch and region datasets with a left join.

```{r}
  # catch_joined <- left_join(catch_cleaned, region_clean, 
  #                           by = c("Region" = "code"))  #use by if the fields are different.
  catch_joined <- left_join(catch_cleaned, region_clean)
  head(catch_joined)
```

#Spread

Make a wide datafram from a long one using the spread function.

```{r}
  catch_wide <- catch_cleaned %>% 
    filter(Year>1990) %>%
    select(-catch_thousands) %>% 
    spread(key=Year, value=catch)
  head(catch_wide) 
```

#Separate and Unite 

Collect some dates.

```{r}
  dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)
  dates_df
```

Separate date information into columns.

```{r}
  dates_sep <- dates_df %>% 
    separate(col=date, into = c("month", "day", "year"), sep = "/", remove="F")
  head(dates_sep)
```

Do some uniting.

```{r}

  dates_unite <- dates_sep %>% 
    unite(date_iso,year,month,day, sep = "-")

  head(dates_unite)

```


