---
title: "index"
author: "Patrick"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=F, message=F, echo=F}
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
```


# Example web page.

## Data Tidying

Read in escapemet data from:  
Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r, results=F}
  esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors=F)
  head(esc)
```

Produce annual regional counts by salmon species.  Create annual data, aggregate into regions.

* `separate` the date into year - month - day
* Use `group_by` and `summarize` to calculate the sum over unique region-year-species combinations.
* `filter` by salmon species.

#Static Plots

```{r}
  annual_esc <- esc %>% 
    separate(sampleDate, into=c("year","month","day"), sep = "-") %>% 
    mutate(year = as.numeric(year)) %>% 
    group_by(SASAP.Region,Species,year) %>% 
    summarize(escapement = sum(DailyCount)) %>% 
    filter(Species %in% c("Chinook","Sockeye","Pink", "Chum", "Coho"))
  head(annual_esc)
```

Create an interactive data data table of the dataframe just generated.
```{r}
  datatable(annual_esc)
```


#Static Plots

Make a bar plot.

```{r}
  ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region)) +
    geom_col()  #<- have to add the type of plot here: a column plot.
    # the aes call maps part of dataframe to part of your plot.
    # see ggplot cheat sheet for more info.
    #geom_col(fill="blue") make the fill blue.
```

Make a timeseries, escapement by species.
```{r}

  #Create a theme object.
  my_theme = theme_bw() + 
    theme(plot.title = element_text(hjust=0.5)) #Make the title centered.

  ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), 
         aes(x = year, y = escapement, color = Species)) +
    geom_line() + 
    geom_point() + #You can add another call to a function with a different aesthetic...
    ylab(bquote("Escapement "~(Wm^-2))) +
    ggtitle("Kodiak Salmon Escapement") + 
    my_theme  #Add a theme object.
  
    # Or add theme directly without creating an object first.
    # theme_bw() + 
    # theme(plot.title = element_text(hjust = 0.5)) #Make title centered.


# bquote('x axis'~(Å^2))
```

Plot timeseries of escapement by species for all regions.
```{r, fig.height=15, fig.width=10}
  p <- ggplot(annual_esc, aes(x=year, y=escapement, color = Species)) +
    geom_line() +
    geom_point() + 
    facet_wrap(~SASAP.Region, scales = "free", ncol=2) +  #Making the multiple plots. 
    my_theme
  #free axes in facet wrap:  free_x, free_y, free

  p  #Print the object p.

  ggsave("images/region_escapement.png", plot=p, height=15, width=10, units="in")
  
  #Grid.arrage another option.
```

```{r}
  write.csv(annual_esc,"derived_data/annual_escapement.csv",
            row.names=F)
```


#Interactive Match


