---
title: "Netcdf testing"
author: "Patrick"
date: "1/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message =F, warning =F}
  library(ncdf4)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
```

#Read in data from a file.

Open the netcdf file.

```{r}
  nc <- nc_open("data/WG2013CTD.nc")
```

Print the netcdf file object.

```{r}
  print(nc)
```

Get the names of all the variables and print out.

```{r}
  vars <- attributes(nc$var)$names
  vars
```

Extract some data from the file.

Get the dimension names of the variable.
```{r}
  dims <- attributes(nc$dim)$names  
```

Get the salinity, depth and time variables.
```{r}
  salinity <- ncvar_get(nc, "sal") #get the salinity data.
  depth <- ncvar_get(nc, "z") #depth
  time_mat <- ncvar_get(nc, "time") #time
```

# Reformat the output.

Convert the matlab datetime numbers into format readable by R.
```{r}
  time <- as.POSIXct((time_mat + 719529)*86400, origin = "1970-01-01", tz = "UTC")
```

Coerce into a data frame.
```{r}
  salinity_data <- as.data.frame(salinity)
```

Assign column names to depth values.
```{r}
  names(salinity_data) <- as.character(depth)
```

Reshape the data and add time.
```{r}
  salinity_data_long <- salinity_data %>% 
    mutate(time = time) %>% 
    gather(key = "depth", value = "salinity", -time) %>% 
    mutate(depth = as.numeric(depth)) %>% 
    arrange(time)
  head(salinity_data_long)
```

#Plots 
Raster plot, option 1.
```{r}
  ggplot(salinity_data_long, mapping = aes(x=time, y=depth, fill = salinity)) +
    geom_raster()
```

Raster plot, option 2.
```{r}
  ggplot(salinity_data_long, mapping = aes(x=salinity, 
                                           y=depth,
                                           color=time,
                                           group = time)) + #lines are grouped by time, not connected.
    geom_line(size=0.1) + 
    scale_y_reverse()
```





